Guarded Recursion in Agda via Sized Types

### 2019-02-03 code-oneclock-paper/latex/Introduction.tex
127-128

We focus on two of these techniques in which productivity is encoded
at the level of types: sized types and guarded recursion. A sized type
 is a type annotated with the number of unfolding that
elements of this type can undergo \cite{Hughes96}. Sized types have been implemented
in Agda and can be used in
combination with coinductive records to specify
coinductive types \cite{Copatterns,AVW-normalization}.  Guarded recursion \cite{Nakano}
is a different approach where the type system is enhanced with a
modality, called ``later'' and written \IC{▻}, encoding time delay in
types. The later modality comes with a general fixpoint combinator for
programming with productive recursive functions and allows the
specification of guarded recursive types, \eg the type of guarded
streams satisfying the isomorphism \F{Str} \Ar{A} \F{≃} \Ar{A} \F{×}
\IC{▻} (\F{Str} \Ar{A}). The tail of a guarded stream is available only
one time step ahead from now, as evidenced by the application of the
later modality to \F{Str} \Ar{A}. Coinductive types can be constructed
from guarded recursive types in type theories that additionally
support clock variables and universal quantification over them
\cite{atkey2013productive,BahrGM17}. Currently there is no
implementation of a calculus for guarded recursion.
%%  nor guarded recursion
%% appears as a selectable feature in a proof assistant.

>>>>

We focus on two of these techniques in which productivity is encoded
at the level of types: sized types and guarded recursion. A sized type
 is a type annotated with the number of unfolding that
elements of this type can undergo \cite{Hughes96}. Sized types have been implemented
in Agda and can be used in
combination with coinductive records to specify
coinductive types \cite{Copatterns,AVW-normalization}, as exemplified
by Abel and Chapman's work \cite{AC14}.  Guarded recursion \cite{Nakano}
is a different approach where the type system is enhanced with a
modality, called ``later'' and written \IC{▻}, encoding time delay in
types. The later modality comes with a general fixpoint combinator for
programming with productive recursive functions and allows the
specification of guarded recursive types, \eg the type of guarded
streams satisfying the isomorphism \F{Str} \Ar{A} \F{≃} \Ar{A} \F{×}
\IC{▻} (\F{Str} \Ar{A}). The tail of a guarded stream is available only
one time step ahead from now, as evidenced by the application of the
later modality to \F{Str} \Ar{A}. Coinductive types can be constructed
from guarded recursive types in type theories that additionally
support clock variables and universal quantification over them
\cite{atkey2013productive,BahrGM17}. Currently there is no
implementation of a calculus for guarded recursion.
%%  nor guarded recursion
%% appears as a selectable feature in a proof assistant.

### 2019-02-03 code-oneclock-paper/latex/paper.tex
127-128

\begin{abstract}
  In type theory, programming and reasoning with possibly
  non-terminating programs and potentially infinite objects is
  achieved with coinductive types. Recursively defined
  programs using such types need to be productive to guarantee consistency, and proof
  assistants such as Agda and Coq use syntactic
  productivity checks for that.
  However, this convolutes programs with coinductive types.

  One way to overcome this issue, is by encoding productivity at
  the level of types, so that the type system forbids the
  implementation of non-productive corecursive programs.
%%, meaning
%%  that although their definitions can be unfolded \emph{ad infinitum},
%%  finite unfoldings are guaranteed to produce . 
  In this paper, we compare two such approaches: guarded recursion and sized types. More specifically,
  we show how to simulate guarded recursion in Agda using sized
  types. We formalize the syntax of a simple type theory for
  guarded recursion, which is a variant of Atkey and McBride's
  type system for productive coprogramming. Then we give a denotational semantics
  using presheaves over the preorder of sizes. 
%Sized types are fundamentally employed in the interpretation of the characteristic features of guarded recursion, notably the fixpoint combinator.
%%extending type theory with
%%  coinduction and coinductive types.  More specifically, we show that
%%  guarded recursion and guarded recursive types can be simulated with
%%  sized types.  We do this by defining a model of guarded recursion in
%%  type theory extended with sized types.  The model is variation of
%%  the topos of trees, and it is formalized in Agda.
\end{abstract}

@@

Finally, we want to understand whether it is possible to reduce sized
types to guarded recursion.  This does not seems directly doable,
since sizes are not uniquely employed with coinductive types as a tool
for easing the productivity check of corecursive definitions, but they
are also used as auxiliary arguments of inductive types for
facilitating the termination check of recursive definitions. It seems
that, in order to capture this dual aspect of sizes, one %needs to
%extend the existing type theories for guarded recursion with
%additional features.
%%
%% Bibliography
%%

>>>>

\begin{abstract}
  In type theory, programming and reasoning with possibly
  non-terminating programs and potentially infinite objects is
  achieved using coinductive types. Recursively defined programs of
  these types are required to be productive to guarantee the
  consistency of the type system. Proof assistants such as Agda and
  Coq traditionally employ strict syntactic productivity checks, which
  commonly make programming with coinductive types convoluted.  One
  way to overcome this issue is by encoding productivity at the level
  of types, so that the type system forbids the implementation of
  non-productive corecursive programs.
%%, meaning
%%  that although their definitions can be unfolded \emph{ad infinitum},
%%  finite unfoldings are guaranteed to produce . 
  In this paper we compare two different approaches to type-based
  productivity: guarded recursion and sized types. More specifically,
  we show how to simulate guarded recursion in Agda using sized
  types. We formalize the syntax of a simple type theory for guarded
  recursion, corresponding to a variant of Atkey and McBride's
  calculus for productive coprogramming
  \cite{atkey2013productive}. Then we give a denotational semantics
  using presheaves over the preorder of sizes. Sized types are
  fundamentally employed in the interpretation of the characteristic
  features of guarded recursion, notably the fixpoint combinator.
%Sized types are fundamentally employed in the interpretation of the characteristic features of guarded recursion, notably the fixpoint combinator.
%%extending type theory with
%%  coinduction and coinductive types.  More specifically, we show that
%%  guarded recursion and guarded recursive types can be simulated with
%%  sized types.  We do this by defining a model of guarded recursion in
%%  type theory extended with sized types.  The model is variation of
%%  the topos of trees, and it is formalized in Agda.
\end{abstract}

@@

Finally, we want to understand whether it is possible to reduce sized
types to guarded recursion.  This does not seems directly doable,
since sizes are not uniquely employed with coinductive types as a tool
for easing the productivity check of corecursive definitions, but they
are also used as auxiliary arguments of inductive types for
facilitating the termination check of recursive definitions. It seems
that, in order to capture this dual aspect of sizes, one needs to
extend the existing type theories for guarded recursion with
additional features.
%%
%% Bibliography
%%
